Forecasting the 2024 US Presidential Election

### 2024-10-30 paper/paper.qmd
added model

\newpage  
# Data {#sec-data}

## Overview

@@

# Function to create the plot with y-axis from 0 to 100 and data labels
create_comparison_plot <- function(plot_data) {
  ggplot(plot_data, aes(
    x = reorder(Variable, -Mean), 
    y = Mean, 
    fill = Group, 
    label = round(Mean, 1)
  )) +
    geom_col(position = position_dodge(width = 0.9)) +
    geom_text(
      position = position_dodge(width = 0.9), 
      vjust = -0.3, 
      size = 3
    ) +
    scale_fill_manual(
      values = c("Blue States" = "blue", "Red States" = "red")
    ) +
    labs(
      title = "Demographic Differences Between Top Red and Blue States",
      x = "Demographic Variable",
      y = "Mean Percentage",
      fill = "State Group"
    ) +
    scale_y_continuous(limits = c(0, 100)) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 7)
    )
}

@@

column_refs <- list(
  democrat = "democrat_pct",
  republican = "republican_pct",
  highly_educated = "What is the highest level of education you have attained? College graduate",
  lower_educated = "What is the highest level of education you have attained? High school or less",
  female = "Can you please tell me your gender? Female",
  male = "Can you please tell me your gender? Male",
  non_binary_other = "Can you please tell me your gender? Nonbinary or other",
  caucasian = "For statistical purposes only, can you please tell me your ethnicity? White or Caucasian",
  hispanic_latino = "For statistical purposes only, can you please tell me your ethnicity? Hispanic or Latino of any race",
  vote_trump = "Who did you vote for in the 2020 election? Donald Trump",
  vote_biden = "Who did you vote for in the 2020 election? Joe Biden",
  biden_approve = "Do you approve or disapprove of the job Joe Biden is doing as President? Approve",
  biden_disapprove = "Do you approve or disapprove of the job Joe Biden is doing as President? Disapprove",
)
```

@@

# Apply a 14-day rolling average (increase smoothing) to each group
time_data <- time_data %>%
  group_by(party) %>%
  mutate(pct_smooth = rollapply(pct, width = 7, FUN = mean, align = "right", fill = NA)) %>%
  ungroup() %>%
  mutate(biden_approval_smooth = rollapply(`Do you approve or disapprove of the job Joe Biden is doing as President? Approve`, 
                                           width = 7, FUN = mean, align = "right", fill = NA))

# Separate data for Democrats, Republicans, and Biden Approval
# Filter data for Democrats where state is NA
democrat_data <- time_data %>%
  filter(party == "DEM" & is.na(state))

# Filter data for Republicans where state is NA
republican_data <- time_data %>%
  filter(party == "REP" & is.na(state))

biden_approval <- time_data %>%
  filter(is.na(state)) %>%
  select(end_date, biden_approval_smooth)

>>>>

\newpage  
# Data {#sec-data}  
## Overview

@@

# Function to create the plot with y-axis from 0 to 100 and data labels
create_comparison_plot <- function(plot_data) {
  ggplot(plot_data, aes(
    x = reorder(Variable, -Mean), 
    y = Mean, 
    fill = Group, 
    label = round(Mean, 1)
  )) +
    geom_col(position = position_dodge(width = 0.9)) +
    geom_text(
      position = position_dodge(width = 0.9), 
      vjust = -0.3, 
      size = 1
    ) +
    scale_fill_manual(
      values = c("Blue States" = "blue", "Red States" = "red")
    ) +
    labs(
      title = "Demographic Differences Between Top Red and Blue States",
      x = "Demographic Variable",
      y = "Mean Percentage",
      fill = "State Group"
    ) +
    scale_y_continuous(limits = c(0, 100)) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 4)
    )
}

@@

column_refs <- list(
  democrat = "democrat_pct",
  republican = "republican_pct",
  highly_educated = "What is the highest level of education you have attained? College graduate",
  lower_educated = "What is the highest level of education you have attained? High school or less",
  female = "Can you please tell me your gender? Female",
  male = "Can you please tell me your gender? Male",
  non_binary_other = "Can you please tell me your gender? Nonbinary or other",
  caucasian = "For statistical purposes only, can you please tell me your ethnicity? White or Caucasian",
  hispanic_latino = "For statistical purposes only, can you please tell me your ethnicity? Hispanic or Latino of any race",
  vote_trump = "Who did you vote for in the 2020 election? Donald Trump",
  vote_biden = "Who did you vote for in the 2020 election? Joe Biden",
  biden_approve = "Do you approve or disapprove of the job Joe Biden is doing as President? Approve",
  biden_disapprove = "Do you approve or disapprove of the job Joe Biden is doing as President? Disapprove",
  previous_vote_trump ="Who did you vote for in the 2020 election? Donald Trump",
  previous_vote_joe ="Who did you vote for in the 2020 election? Joe Biden"

)
```

@@

# Average entries for the same date and party, then apply a 14-day rolling average for smoothing
time_data <- time_data %>%
  group_by(end_date, party) %>%
  summarise(
    pct = mean(pct, na.rm = TRUE),
    biden_approval = mean(`Do you approve or disapprove of the job Joe Biden is doing as President? Approve`, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  group_by(party) %>%
  mutate(
    pct_smooth = rollapply(pct, width = 6, FUN = mean, align = "right", fill = NA)
  ) %>%
  ungroup() %>%
  mutate(
    biden_approval_smooth = rollapply(biden_approval, width = 26, FUN = mean, align = "right", fill = NA)
  )

# Separate data for Democrats, Republicans, and Biden Approval
# Filter data for Democrats where state is NA
democrat_data <- time_data %>%
  filter(party == "DEM")

# Filter data for Republicans where state is NA
republican_data <- time_data %>%
  filter(party == "REP")

biden_approval <- time_data %>%
  filter(party == "DEM" | party == "REP") %>%
  select(end_date, biden_approval_smooth) %>%
  distinct()
